p-NLI: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-21 paper/intro.tex
update structure

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{} on interpretation of convolution neural network (CNN) model from both machine learning and visualization community has reveal the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know regarding the inner mechanism of these deep neural network models. 
%
Many of the existing techniques study the model as an invariant object, where the model's behavior are recorded and analyzed in an offline fashion.
%
However, the exploration natural of these type of analysis often lead to \emph{what if ...} types of questions. What if I perturbed the current input? Will the prediction be stable? What if when change the internal state of the model, how could the change affect the prediction. What if the current prediction is wrong, how and where should we change the model in order to produce the correct result? These type of dynamic queries is a natural way for us to gain understanding and come up with hypothesis of the model mechanisms by \emph{interrogating} the relationship between the input, internal states, and output prediction. Yet we do not often need to setup extra specialized experiment to answer.

Compare to image classification task, natural language processing systems have a number of innate restriction (e.g., the discrete nature of words) that makes interpreting them even more challenging. For example, \emph{feature visualization}~\cite{}, an often deployed technique for illustrating what kind of image feature a given part of the network (e.g., neuron, layer, or channel) captures, can not be easily generalized and applied for natural language due to the requirement of a continuous solution space (i.e., the image pixel values). Even though we can encode a word as an vector~\cite{}, the word embedding space is still a discretely defined space, where the interpolation between two vectors (i.e., two words) do not corresponds to explicit meaning.

On the one hand, recognize where the model fail is non-trivial. The standard evaluation approach (i.e., the prediction accuracy on the test set) provides limited information and cannot answer many important questions (e.g., how stable the prediction is, which part of the model may contribute to the failure) that are essential to the model diagnostic process. On the other hand, infer the cause of model failure is even more challenging. Neural network models are often referred to as black boxes due to the hard to recognize  difficult to interpret parameter, hypothesizing what happened in prediction process require the ability to peek inside the model and understand the intricate relationships among model inputs, critical internal mechanisms, as well the output predictions.

@@

Questions such as, how stable predictions are concerning small perturbation in the input, how are these predictions arrived (i.e., the model may come up with the correct prediction by pickup unintended feature/alignment), and what are the often made mistakes, are all essential in systematic examination and evaluation of the model. 
%%%%%%%%% accessing internal states are important %%%%%%%%%%%%

% Making sense and explaining predictions made by neural networks is also becoming a necessity.
%
Many works~\cite{RibeiroSinghGuestrin2016} have been dedicated to providing intuitive explanations for a given prediction, several of which approach the problem from a model agnostic approach that made them applicable to different applications (i.e., by fitting a simplier linear model near the prediction of interests).
%
Despite being invaluable for the providing human understandable explaination, the lack of the ability to access and explore the internal states of the model, which is vitally important to make sense of why a model fail, limits their ability for more in depth exploration and evaluation of the model.

%%%%%%%%%% interrogate relationship between different components, modified different components %%%%%%%%%%%
Beside looking into the internal states, understanding how each of the component of the model interact with input and output as well as with each other is the crucial for truly exmaine the mechanism of the model.
%
Look at how the model work in action and interrogate relationship between different component, i.e., how the change made to one part of model affect the other pieces present, is the key to gain the full picture.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{Language Inference Model Pipeline.}
\label{fig:projTransition}
\end{figure}

In this work, we introduce a visualization system that through a tight yet flexible integration between visual analytic elements and the underlying model, allow a user to interrogate the model by perturbing the input, the internal state, as well as prediction while observing changes in other parts of the pipeline.
We use natural language inference task~\cite{BowmanAngeliPotts2015} as an example to illustrate how this system help natural language processing (NLP) researchers quickly identify the potential limitation of an NLP model, probe the inner states of the model for interpreting key mechanism such as attention, and ultimately come up with hypothesis on why the model fails.

@@

>>>>

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{} on interpretation of convolution neural network (CNN) model from both machine learning and visualization community has reveal the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know regarding the inner mechanism of these deep neural network models.
%
Compare to image classification task, natural language processing systems have a number of innate restriction (e.g., the discrete nature of words) that makes interpreting them even more challenging. For example, \emph{feature visualization}~\cite{}, an often deployed technique for illustrating what kind of image feature a given part of the network (e.g., neuron, layer, or channel) captures, can not be easily generalized and applied for natural language due to the requirement of a continuous solution space (i.e., the image pixel values). Even though we can encode a word as an vector~\cite{}, the word embedding space is still a discretely defined space, where the interpolation between two vectors (i.e., two words) do not corresponds to explicit meaning.
%
In this work, we introduce a visualization system specific tailor to natural language model, specifical natural language inference, in which we address many domain specific challenges.

Many of the existing techniques study the model as an invariant object, where the model's behavior are recorded and analyzed in an offline fashion.
%
However, the exploration natural of these type of analysis often lead to \emph{what if ...} types of questions. What if I perturbed the current input? Will the prediction be stable? What if we change one of the key internal states of the model, how would the change affect the prediction. What if the current prediction is wrong, how and where should we change the model in order to produce the correct result? These type of dynamic queries is a natural way for us to gain understanding and come up with hypothesis of the model mechanisms by \emph{interrogating} the relationship between the input, internal states, and output prediction. Yet, we often need to tediously setup specialized experiment to obtain the answers.
%
In this work, we aim to provide quick and intuitive answers to these \emph{what if ...} questions. Instead of viewing the model as an invariant entitle, we employ a perturbation centric mentally to probe the model and examine how changes in one part of the pipeline affect others, which in term provide a new perspective to look at the model interpretation problem.

@@

%Questions such as, how stable predictions are concerning small perturbation in the input, how are these predictions arrived (i.e., the model may come up with the correct prediction by pickup unintended feature/alignment), and what are the often made mistakes, are all essential in systematic examination and evaluation of the model.
%%%%%%%%% accessing internal states are important %%%%%%%%%%%%

%%%%%%%%%% interrogate relationship between different components, modified different components %%%%%%%%%%%

@@

% On the one hand, recognize where the model fail is non-trivial. The standard evaluation approach (i.e., the prediction accuracy on the test set) provides limited information and cannot answer many important questions (e.g., how stable the prediction is, which part of the model may contribute to the failure) that are essential to the model diagnostic process. On the other hand, infer the cause of model failure is even more challenging. Neural network models are often referred to as black boxes due to the hard to recognize  difficult to interpret parameter, hypothesizing what happened in prediction process require the ability to peek inside the model and understand the intricate relationships among model inputs, critical internal mechanisms, as well the output predictions.

### 2018-03-21 paper/paper_entailVis.tex
update intro
update structure

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}
%% please declare the paper type of your paper to help reviewers, only shown in review mode
%% choices:
%% * algorithm/technique
%% * application/design study
%% * evaluation
%% * system
%% * theory/model
\vgtcpapertype{application}

%% Paper title.
\title{Where and How Do Models Fail? A Visual Exploration Tool for Analyzing and Interpreting Natural Language Inference}
%% This is how authors are specified in the journal style

@@

%other entries to be set up for journal
\shortauthortitle{Text Entailment Model Visualization}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
% What are we doing in this paper
\abstract{
With the recent advances in deep learning, neural network based models have obtained the state-of-the-art performance for many linguistic tasks in natural language processing. The disruptive progress also brings enormous challenges.
The opaque nature of neural network model leads to hard to debug system and difficult to interpret mechanisms. In this work, we use natural language inference task as an example to illustrate how a flexible visualization system can not only help NLP researchers quickly identify the potential limitation of a model, but also facilitate the interpretation the model mechanisms by allowing interactive exploration of the interplay between model input, model inner states, and the prediction.
} % end of abstract

@@

%% the only exception to this rule is the \firstsection command
\input{intro.tex}
\input{relatedWorks.tex}
\input{background.tex}
% \input{tasks.tex}
\input{perturbation.tex} % how to do the perturbation
\input{attention.tex} % how to encode the information
\input{pipeline.tex}
\input{application.tex}
%%%%%%%%%%%%% identify errors %%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% interpret attention %%%%%%%%%%%%%%%%%%%

>>>>

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{1095}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}
%% please declare the paper type of your paper to help reviewers, only shown in review mode
%% choices:
%% * algorithm/technique
%% * application/design study
%% * evaluation
%% * system
%% * theory/model
\vgtcpapertype{Application}

%% Paper title.
\title{
p-NLI: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models
%Where and How Do Models Fail? A Visual Exploration Tool for Analyzing and Interpreting Natural Language Inference Models
}
%% This is how authors are specified in the journal style

@@

%other entries to be set up for journal
\shortauthortitle{Natural Language Inference Model Visualization}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
% What are we doing in this paper
\abstract{
With the recent advances in deep learning, neural network based models have obtained the state-of-the-art performance for many linguistic tasks in natural language processing. The disruptive progress also brings enormous challenges.
The opaque nature of neural network model leads to hard to debug system and difficult to interpret mechanisms. 
%In this work, we use natural language inference task as an example to illustrate how a flexible visualization system can not only help NLP researchers quickly identify the potential limitation of a model, but also facilitate the interpretation the model mechanisms by allowing interactive exploration of the interplay between model input, model inner states, and the prediction.
In this work, we introduce a visualization system that through a tight yet flexible integration between visualization elements and the underlying model, allow a user to interrogate the model by perturbing the input, the internal state, as well as prediction while observing changes in other parts of the pipeline.
%
We use natural language inference task~\cite{BowmanAngeliPotts2015} as an example to illustrate how this system help natural language processing (NLP) researchers quickly identify the potential limitation of an NLP model, probe the inner states of the model for interpreting key mechanism such as attention, and ultimately come up with hypothesis on why the model fails.
} % end of abstract

@@

%% the only exception to this rule is the \firstsection command
\input{intro.tex}
\input{relatedWorks.tex}

\input{background.tex}
%\input{tasks.tex}
\input{pipeline.tex}
\input{perturbation.tex} % how to do the perturbation
\input{attention.tex} % how to encode the information

\input{application.tex}
%%%%%%%%%%%%% identify errors %%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% interpret attention %%%%%%%%%%%%%%%%%%%

### 2018-03-21 paper/relatedWorks.tex
update structure

How to interpret classification task
\begin{itemize}
    \item What feature / dimension / part-of-the-input should change to
    alter the outcome
    \item What Examples can you provide that is similar to the current
    instance
\end{itemize}
\begin{itemize}
    \item Interacting with Predictions: Visual Inspection of Black-box Machine Learning Models,
    Use partial dependency (varying along one feature and observer how prediction changes)
    \item “Why Should I Trust You?” Explaining the Predictions of Any Classifier (LIME)
\end{itemize}

>>>>

How to interpret classification task
\begin{itemize}
    \item What feature / dimension / part-of-the-input should change to
    alter the outcome
    \item What Examples can you provide that is similar to the current
    instance
\end{itemize}
\begin{itemize}
    \item Interacting with Predictions: Visual Inspection of Black-box Machine Learning Models,
    Use partial dependency (varying along one feature and observer how prediction changes)
    \item “Why Should I Trust You?” Explaining the Predictions of Any Classifier (LIME)
\end{itemize}
% Making sense and explaining predictions made by neural networks is also becoming a necessity.
%
Many works~\cite{RibeiroSinghGuestrin2016} have been dedicated to providing intuitive explanations for a given prediction, several of which approach the problem from a model agnostic approach that made them applicable to different applications (i.e., by fitting a simplier linear model near the prediction of interests).
%
Despite being invaluable for the providing human understandable explanation, the lack of the ability to access and explore the internal states of the model, which is vitally important to make sense of why a model fail, limits their ability for more in depth exploration and evaluation of the model.
